A user-extensible and adaptable parser architecture

### 2008-05-27 papers/sgai-2008/paper.tex
Added lots of content from logparser.pdf to the results section, including lots of graphs.  I'll soon have to cut content out, but I'll leave it in for now and ask Carl's advice.

    \item [regex] The \regex{} to match the log line against.  The \regex{}
        will first have several keywords expanded: this simplifies reading
        and writing rules; avoids needless repetition of complex \regex{}
        components; allows the components to be corrected and/or improved
        in one location; and makes each \regex{} largely self-documenting.
        For efficiency the keywords are expanded and every rule's \regex{}
        is compiled before attempting to parse the log file --- otherwise
        each \regex{} would be recompiled each time it was used, resulting
        in a large, data dependent slowdown.  Rule efficiency concerns are
        discussed in \sectionref{rule efficiency}, with the impact of
        compiling and caching \regexes{} covered in \sectionref{Caching
        each regex}.

@@

>>>>

    \item [regex] The \regex{} to match the log line against.  The \regex{}
        will first have several keywords expanded: this simplifies reading
        and writing rules; avoids needless repetition of complex \regex{}
        components; allows the components to be corrected and/or improved
        in one location; and makes each \regex{} largely self-documenting.
        For efficiency the keywords are expanded and every rule's \regex{}
        is compiled before attempting to parse the log file --- otherwise
        each \regex{} would be recompiled each time it was used, resulting
        in a large, data dependent slowdown.  Rule efficiency concerns are
        discussed in \sectionref{rule efficiency}.

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
deal with 75 MB log files containing 300,000 log lines (generated daily on
a mail server handling mail for approximately 700 users --- large scale
mail servers would have much larger log files on a daily basis).  When
generating the data for the graphs included in this document,
\numberOFlogFILES{} log files (totaling 10.08 GB, \numberOFlogLINEShuman{}
log lines) were each parsed 10 times, the first run discarded, and the
execution time for the remaining 9 runs averaged.  The first run is
discarded for two reasons:

\begin{enumerate}

    \item The execution time will be higher because the log file must be
        read from disk, whereas for subsequent runs the log file will be
        cached in memory by the operating system.

    \item The execution time will also be higher because the rule ordering
        will be sub-optimal compared to subsequent runs.

\end{enumerate}

Saving results to the database was disabled for the test runs, as that
dominates the run time of the program, and the tests are aimed at measuring
the speed of the parser rather than the speed of the database and the disks
the database is stored on.

\subsection{Architecture Scalability}

MINIMUM VERSUS MAXIMUM NUMBER OF RULES\@.

An important property of a parser is how execution time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
Figure~\refwithpage{execution time vs file size vs number of lines graph}
shows the execution time in seconds, file size in MB and tens of thousands
of log lines per log file.  All three lines run roughly in parallel, giving
a visual impression that the algorithm scales linearly with input size.
This impression is borne out by fig.~\refwithpage{execution time vs file
size vs number lines factor} which plots the ratio of file size vs
execution time and ratio of number of log lines vs execution time (higher
is better).  As the reader can see the ratios are quite tightly banded,
showing that the algorithm scales linearly: the much larger log files
between points 60 and 70 on the X axis in fig.~\refwithpage{execution time
vs file size vs number of lines graph} actually cause the ratio to
increase, rather than decrease.  The strange behaviour where larger log
files are parsed more efficiently is explained fully in \sectionref{Why are
there dips in the graphs?}.

\showgraph{build/plot-normal-filesize-numlines}{Execution time vs file
size vs number of lines}{execution time vs file size vs number of lines
graph}

\showgraph{build/plot-normal-filesize-numlines-factor}{Ratio of file
size and number of lines to execution time}{execution time vs file size vs
number lines factor}

\subsection{Rule ordering}

\label{Rule ordering}
\label{rule ordering for efficiency}
\label{rule efficiency}

At the time of writing there are \numberOFrules{} different rules, with the
top 10\% matching the vast majority of the log lines, and the remaining log
lines split across the other 90\% of the rules (as shown in
fig.~\refwithpage{rule hits graph}).  \showgraph{build/plot-hits}{Hits per
rule}{rule hits graph} Assuming that the distribution of log lines is
reasonably steady over time, program efficiency should benefit from trying
more frequently matching rules before those which match less frequently.
To test this hypothesis three full test runs were performed with different
rule orderings:

    \item [normal]  The most optimal order, according to the hypothesis:
        rules which match most often will be tried first.

    \item [shuffle] Random ordering --- the rules will be shuffled once
        before use and will retain that ordering for the entirety of the
        log file.  Note that the ordering will change every time the parser
        is executed, so 10 different orderings will be generated for each
        log file in the test run.  This is intended to represent an
        unsorted rule set.

    \item [reverse] Hypothetically the worst order: the most frequently
        matching rules will be tried last.

Figure~\refwithpage{percentage increase of shuffled and reversed over
normal} shows the percentage increase of execution times.  Overall this
provides a modest but worthwhile performance increase of over 10\%, for a
small investment in time and programming.
\showgraph{build/plot-normal-shuffle-reverse-factor}{Percentage increase of
shuffled and reversed over normal}{percentage increase of shuffled and
reversed over normal}

\subsubsection{Why are there dips in the graphs?}
\label{Why are there dips in the graphs?}

The dips at log files 22 and 62--68 correspond to peaks in log file size in
graph~\refwithpage{execution time vs file size vs number of lines graph},
and peaks in graph~\refwithpage{execution time vs file size vs number lines
factor} (where a peak means that more lines are processed per second, i.e.\
performance is better).  The explanation for this took some time to arrive
at, but it turns out to be reasonably simple.  The large log files in
question were caused by a mail forwarding loop, where the distribution of
log lines is quite different to normal, resulting in different performance
characteristics.  Fewer

The vast majority of log lines when a mail loop occurs are from Postfix
components which have a small number of rules associated with them, whereas
in general \daemon{smtpd} adds the majority of log lines, and also has the
highest number of rules.  \daemon{smtpd} log lines are distributed across
rules much more evenly than the log lines which occur frequently during a
mail loop, so the average number of rules
required to parse a \daemon{smtpd} log line is much higher that the average
number required to parse other log lines.

These two characteristics combine to reduce the average number or rules
required to parse a log line when there is a mail loop, as shown by the
peaks in fig.~\refwithpage{execution time vs file size vs number lines
factor}.  When the rule ordering is reversed the majority of log lines
generated by a mail loop will be parsed with very few rules, whereas
without a mail loop the majority of log lines require a large number of
rules; this leads to a noticeable drop in the average time required to
parse a log line, as shown in fig.~\refwithpage{percentage increase of
shuffled and reversed over normal}.  The number of rules which need to be
consulted when the ordering is shuffled varies between the optimum and
nadir, and the performance varies proportionally.

\subsection{Coverage}
